---
title: "KNZ_L0_data_wrangling"
authors: Smriti Pehim Limbu, Joshua Ajowele
Collaborators: Grassland Rocks working group
date: "2023-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##package upload
```{r}
install.packages(pacman)
library(pacman)
pacman::p_load(readr, ggplot2, dplyr, tidyverse, gtools, lubridate, codyn) #lubridate is to change the date format

```

##set working directory
```{r}
L0_dir <- Sys.getenv("L0DIR")
L1_dir <- Sys.getenv("L1DIR")
L2_dir <- Sys.getenv("L2DIR")
list.files(L0_dir)
```

##Load files, more needs to be loaded

```{r}
df_A1_KNZ <- read.csv(file.path(L0_dir, "KNZ/A1_CME011_Sppcomp.csv"), stringsAsFactors = FALSE)
df_A2_KNZ <- read.csv(file.path(L0_dir, "KNZ/A2_CME012_Biomass.csv"), stringsAsFactors = FALSE)

#Joshua's directory
df_E1_KNZ<-read.csv("C:/Users/JAAJOWELE/Downloads/E1_PPL011_Sppcomp.csv")
df_E2_KNZ<-read.csv("C:/Users/JAAJOWELE/Downloads/E2_PPL012_Biomass.csv")


df_J1_KNZ <- read.csv(file.path(L0_dir, "KNZ/J1_CEE012_Sppcomp.csv"), stringsAsFactors = FALSE)
df_J2_KNZ <- read.csv(file.path(L0_dir, "KNZ/J2_CEE014_Biomass.csv"), stringsAsFactors = FALSE)

#Alex's directory
df_C1_KNZ <- read.csv("C:\\Users\\jasig\\Downloads\\C1_WAT012_Sppcomp.csv")
df_C2_KNZ <- read.csv("C:\\Users\\jasig\\Downloads\\C2_WAT013_Biomass.csv")
df_ppt <- read.csv("C:\\Users\\jasig\\Downloads\\Precipitation_1982_2023.csv")
#(Alex is dumb & is taking this approach) Convert the date column to Date type 
df_ppt$RecDate<- as.Date(df_ppt$RecDate, format = "%m/%d/%Y")
#Extract the year from the date
df_ppt$year <- year(df_ppt$RecDate)
#Reading temp data
df_temp <- read.csv("C:\\Users\\jasig\\Downloads\\Temperature_1982_2023.csv")
```

#climate data

```{r}
df_ppt <- read.csv(file.path(L0_dir, "KNZ/Precipitation_1982_2023.csv"), stringsAsFactors = FALSE)
df_temp <- read.csv(file.path(L0_dir, "KNZ/Temperature_1982_2023.csv"), stringsAsFactors = FALSE)
# Convert the date column to Date type 
df_ppt$RecDate<- as.Date(df_ppt$RecDate, format = "%m/%d/%Y")
# Extract the year from the date
df_ppt$year <- year(df_ppt$RecDate)

# Summarizing temperature data
# Averaging across months
df_temp_ave_month <- df_temp %>% 
  group_by(RECYEAR, RECMONTH) %>%
  summarize(mean=mean(as.numeric(TAVE)))

# Averaging across years
df_temp_ave_year <- df_temp %>% 
  group_by(RECYEAR) %>%
  summarize(mean=mean(as.numeric(TAVE)))
```


#Cleaning CEE data J1 and J2 - Smriti

```{r}

#renaming variables

df_J1_KNZ.clean <- df_J1_KNZ %>% rename(year = Year, species = Spnum2, abundance=Cover)

#adding site name, watershed name, blocks, and higher_order_organization

df_J2_KNZ.clean <- df_J2_KNZ %>% rename(year = Year)  %>% mutate(site = "KNZ") %>% mutate(watershed="HQ") %>% mutate(Block =case_when(Plot%in% c(101, 102, 103, 104, 105, 106, 107, 108, 109, 110)~"1", Plot%in% c(101, 102, 103, 104, 105, 106, 107, 108, 109, 110)~"1",Plot%in% c(201, 202,203, 204, 205, 206, 207, 208, 209, 210)~"2", Plot%in% c(301, 302, 303, 304, 305, 306, 307, 308, 309, 310)~"3",Plot%in% c(401, 402, 403, 404, 405, 406, 407, 408, 409, 410)~"4")) %>%  unite(higher_order_organization,site, watershed,Block,Plot, sep ="_", remove=FALSE)


#calculate richness and evenness

df_J1_richness_evenness <- community_structure(df_J1_KNZ.clean, time.var = "year", replicate.var = "Plot", abundance.var = "abundance", metric="SimpsonEvenness")

#add richness and evenness to the biomass data

df_J2_KNZ.clean.richness.evenness <- left_join (df_J2_KNZ.clean, df_J1_richness_evenness, by = c("year", "Plot"))


```

#cleaning data E1 and E2 - Joshua
```{r}
#check each column for uniqueness for E1 and E2 data
unique(df_E1_KNZ$RecYear)#2020 missing due to covid
unique(df_E2_KNZ$RecYear)#2020 data present
unique(df_E1_KNZ$Treatment)#looks good, 8 treatments
unique(df_E2_KNZ$Treatment)#looks good, 8 treatment
unique(df_E1_KNZ$PlotID)#48 plots
unique(df_E2_KNZ$PlotID)#48 plots

#clean and wrangle sp comp data
df_E1_KNZ_clean<-df_E1_KNZ%>%
  rename(year=RecYear)%>%
  unite(species_name,Genus,Species, sep = "_")%>%
#unique(df_E1_KNZ_clean$species_name)#no obvious duplicate 
  group_by(year, PlotID, Treatment)%>%
  mutate(total_abundance=sum(Abundance))%>%
  group_by(year, PlotID, Treatment, species_name)%>%
  mutate(relative_abundance=(Abundance/total_abundance)*100)%>%
  ungroup()%>%
  #create unique identifier for codyn package for community metric
  unite(plot, PlotID, Treatment, sep="_")
  #ready for richness/diversity calculation with codyn package

#calculate diversity/richness/evenness

E1_richness<-df_E1_KNZ_clean%>%
  community_diversity(time.var="year",abundance.var="relative_abundance",
                      replicate.var="plot",metric = "Shannon")
E1_diversity<-df_E1_KNZ_clean%>%
  #calculates richness and evenness as Evar
  community_structure(time.var="year",abundance.var="relative_abundance",
                      replicate.var="plot")%>%
  left_join(E1_richness, by=c("year","plot"))%>%
  separate_wider_delim(plot,"_", names=c("PlotID","Treatment"))
#convert to integer to allow joining with biomass data
E1_diversity$PlotID=as.integer(E1_diversity$PlotID)

#clean and wrangle complementary biomass data and join with spcomp data

#create a treatment key
treatment_key=tibble(Treatment=levels(factor(df_E2_KNZ$Treatment)),                        treatment_key=c("control","N0P2.5","N0P5","N0P10",
                                    "N10P0","N10P2.5","N10P5","N10P10"))

#clean biomass data and join with spcomp data
df_E1_E2_KNZ_clean<-df_E2_KNZ%>%
  rename(year=RecYear, plot_biomass=ANPP)%>%
  mutate(site="KNZ", fire_frequency=2, grazing="ungrazed", 
         watershed="2c")%>%
  left_join(E1_diversity, by=c("year","Treatment", "PlotID"))%>%
  left_join(treatment_key, by="Treatment")
#ready to merge with other data set

```

#Cleaning data C1 & C2 - Alex
```{r}
##Species comp
#Merging Genus & Species columns
df_C1_KNZ$species <- paste(df_C1_KNZ$Genus,df_C1_KNZ$Species)
#Removing unnecessary columns
df_C1_KNZ = df_C1_KNZ %>%
  select(-c("Rectype","Genus","Species"))
#Renaming columns
df_C1_KNZ.clean <- df_C1_KNZ %>% rename(year = RecYear, treatment = Trt, transect = Transect, plot = Plot, species_code = Spcode, abundance=Cover)
#Adding columns
df_C1_KNZ.clean <- df_C1_KNZ.clean %>% 
  mutate(site = "KNZ")
#Checking columns for uniqueness
unique(df_C1_KNZ.clean$year) #1991-2022
unique(df_C1_KNZ.clean$treatment) #Three treatments (a & i, c starts in 2018)
unique(df_C1_KNZ.clean$plot) #30 plots (no #9?, added plots in 1993)
unique(df_C1_KNZ.clean$transect) #Four transects
## Must consider treatment reversal for this experiment ##

#clean and wrangle spp comp data
df_C1_KNZ.clean<-df_C1_KNZ.clean%>%
  group_by(year, plot, treatment)%>%
  mutate(total_abundance=sum(abundance))%>%
  group_by(year, plot, treatment, species)%>%
  mutate(relative_abundance=(abundance/total_abundance)*100)%>%
  ungroup()%>%
  #create unique identifier for codyn package for community metric
  unite(plot_trt, plot, treatment, sep="_", remove=FALSE)

#calculate diversity/richness/evenness
C1_richness<-df_C1_KNZ.clean%>%
  community_diversity(time.var="year",abundance.var="relative_abundance",
                      replicate.var="plot_trt",metric = "Shannon")
C1_diversity<-df_C1_KNZ.clean%>%
  #calculates richness and evenness as Evar
  community_structure(time.var="year",abundance.var="relative_abundance",
                      replicate.var="plot_trt")%>%
  left_join(C1_richness, by=c("year","plot_trt"))%>%
  separate_wider_delim(plot_trt,"_", names=c("plot","treatment"))
#convert to integer to allow joining with biomass data
C1_diversity$plot=as.integer(C1_diversity$plot)

#Adding in precip data
df_ppt$ppt = as.numeric(df_ppt$ppt)
df_ppt[is.na(df_ppt)] <- 0
df_ppt = df_ppt %>% 
  group_by(year) %>%
  mutate(annualprecip = sum(ppt))
df_ppt = df_ppt %>% 
  distinct(year, annualprecip, .keep_all = TRUE)
df_ppt = df_ppt[-c(1:6)]

df_C1_KNZ.cleanest <- df_C1_KNZ.clean %>%
right_join(df_ppt, by=c("year")) %>%
  mutate(method="sorted_to_species",
         experiment="WAT01",
         nutrients_added = "no_fertilizer",
         nitrogen_added = NA,
         disturbance = "undisturbed",
         grazing = "ungrazed",
         fire_frequency = 1,
         time_since_fire = NA,
         diversity_manipulated = "naturally_assembled",
         measurement_scale_biomass = NA,
         richness = NA,
         evenness = NA,
         measurement_scale_cover = NA) %>%
  unite(higher_order_organization,site,experiment,plot_trt, sep ="_", remove=FALSE)

#Writing sppcomp dataset to upload to L1 folder
write.csv(df_C1_KNZ.cleanest,"C:\\Users\\jasig\\Downloads\\df_C1_KNZ.cleanest.csv", row.names=FALSE)

## Species comp data is clean ##
## There are issues with naming conventions, such as transect and treatment being interchanged between datasets


##Biomass data
df_C2_KNZ.cleaner <- df_C2_KNZ %>%
  select(-c("RECMONTH","CURRENTDEAD","COMMENTS")) %>%
  rename(DataCode=DATACODE, year=RECYEAR, watershed=WATERSHED, treatment=TRANSECT, plot=PLOT, replicate=REPLICATE, livegrass=LIVEGRASS, forbs=FORBS, woody=WOODY)
#Creating ANPP column
df_C2_KNZ.cleaner$livegrass = as.numeric(df_C2_KNZ.cleaner$livegrass)
df_C2_KNZ.cleaner$forbs = as.numeric(df_C2_KNZ.cleaner$forbs)
df_C2_KNZ.cleaner$woody = as.numeric(df_C2_KNZ.cleaner$woody)
df_C2_KNZ.cleaner$total_biomass = df_C2_KNZ.cleaner$livegrass + df_C2_KNZ.cleaner$forbs + df_C2_KNZ.cleaner$woody

unique(df_C2_KNZ.cleaner$treatment) 

#Calculating plot_biomass and more
df_C2_KNZ.clean <- df_C2_KNZ.cleaner %>%
  group_by(year, plot) %>%
  mutate(plot_biomass = sum(total_biomass)) %>%
  mutate(mean_livegrass_rep = mean(livegrass)) %>%
  mutate(mean_forbs_rep = mean(forbs)) %>%
  mutate(mean_woody_rep = mean(woody)) %>%
  mutate(mean_biomass_rep = mean(total_biomass)) %>%
  mutate(site = "KNZ") %>%
#create unique identifier to consolidate reps
  unite(plot_trt, plot, treatment, sep="_", remove=FALSE)

#Subsetting to just include plots and their average biomass per functional group (to account for four reps each)
df_C2_KNZ.noreps <- df_C2_KNZ.clean %>% 
  distinct(year, plot_trt, plot_biomass, .keep_all = T)
df_C2_KNZ.noreps <- df_C2_KNZ.noreps[-c(7:11)] %>%
  unite(higher_order_organization,site,plot_trt, sep ="_", remove=FALSE)

#Adding precip data
df_C2_KNZ.cleanest <- df_C2_KNZ.noreps %>%
right_join(df_ppt, by=c("year")) %>%
  mutate(experiment="WAT01",
         nutrients_added = "no_fertilizer",
         nitrogen_added = NA,
         disturbance = "undisturbed",
         grazing = "ungrazed",
         fire_frequency = 1,
         time_since_fire = NA,
         diversity_manipulated = "naturally_assembled",
         measurement_scale_biomass = NA,
         richness = NA,
         evenness = NA,
         measurement_scale_cover = NA) %>%
  unite(higher_order_organization,site,experiment,plot_trt, sep ="_", remove=FALSE)

unique(df_C2_KNZ.cleanest$treatment) #6 unique treatment (i, c, ni, b, d, a)

#Writing biomass dataset to upload to L1 folder
write.csv(df_C2_KNZ.cleanest,"C:\\Users\\jasig\\Downloads\\df_C2_KNZ.cleanest.csv", row.names=FALSE)


#Joining dataframes (Not ready yet)
df_C1_C2_KNZ_clean<-df_C2_KNZ.noreps%>%
  left_join(C1_diversity, by=c("year","plot"))

## Running into issues with naming, would like to chat at next Konza working hours ##

```


# Cleaning data D1 and D2 - Matt
```{r}

## bring in RaMPs data
# ramps <- read.csv("D1_D2_RMP011_Sppcomp_Biomass .csv",header = TRUE, sep = ",") # NOTICE there is a space before the .csv
#Note from Maya: I am going to read in data from the shared Google drive instead

ramps <- read.csv(file.path(L0_dir, "KNZ/D1_D2_RMP011_Sppcomp_Biomass .csv"), header = TRUE, sep = ",", stringsAsFactors = FALSE) # NOTICE there is a space before the .csv

ramps$RecYear <- as.factor(ramps$RecYear)
ramps$season <- as.factor(ramps$season)
ramps$RampNo <- as.factor(ramps$RampNo)
ramps$Species <- as.factor(ramps$Species)


##### ANPP #####
RaMPs_ANPP_working <- ramps %>% 
  mutate(A_g_m2 = A*10,
         B_g_m2 = B*10,
         C_g_m2 = C*10,
         D_g_m2 = D*10) %>% 
  group_by(RecYear, season, RampNo) %>% 
  summarize(A_ANPP = sum(A_g_m2),
            B_ANPP = sum(B_g_m2),
            C_ANPP = sum(C_g_m2),
            D_ANPP = sum(D_g_m2)) %>% 
  tidyr::gather(key="plot", value="plot_biomass", A_ANPP, B_ANPP, C_ANPP, D_ANPP) %>% 
  mutate(plot = plyr::revalue(plot, c("A_ANPP" = "A",
                                      "B_ANPP" = "B",
                                      "C_ANPP" = "C",
                                      "D_ANPP" = "D")))

ambient_ramps <- c(3, 4, 6, 7, 11, 15)
altered_ramps <- c(2,5,9, 10, 12, 13)
control_ramps <- c(1, 8, 14)

levels(as.factor(RaMPs_ANPP_working$RampNo))
class(RaMPs_ANPP_working$RampNo)

RaMPs_ANPP <- RaMPs_ANPP_working %>% 
  dplyr::rename(year = RecYear,
                subplot = plot) %>% 
  mutate(site = "KNZ",
         ramps = "RampNo",
         experiment = "RaMPs",
         treatment  = case_when(RampNo %in% ambient_ramps ~ "ambient",
                                RampNo %in% altered_ramps ~ "altered",
                                RampNo %in% control_ramps ~ "control"),
         block = case_when(RampNo %in% c(1:5) ~ "Block1",
                           RampNo %in% c(6:10) ~ "Block2",
                           RampNo %in% c(11:15) ~ "Block3")) %>%
  filter(year != "1997") %>% # 1997 was year before shelters were put up
  #group_by(year, season, RampNo, site, ramps, experiment, treatment) %>% 
  #summarize(plot_biomass = mean(plot_biomass)) %>% 
  tidyr::unite("higher_order_organization", c(experiment, block), remove = F) %>% #ramps, RampNo
  tidyr::unite("uniqueID", c(higher_order_organization, ramps, RampNo, subplot), remove = F) %>%
  dplyr::rename(plot = RampNo) %>% 
  dplyr::select(year, season, site, plot, higher_order_organization, uniqueID, treatment, plot_biomass) %>% 
  mutate(nutrients_added = "no_fertilizer",
         nitrogen_added = 0,
         disturbance = "undisturbed",
         grazing = "ungrazed",
         fire_frequency = 1,
         time_since_fire = NA,
         diversity_manipulated = "naturally_assembled",
         measurement_scale_biomass = NA,
         richness = NA,
         evenness = NA,
         measurement_scale_cover = NA) %>% 
  mutate(treatment_comment = case_when(treatment == "altered" & as.numeric(year) <= 2001 ~ "Could be reduced quantity, altered pattern, or both",
                                       treatment == "altered" & as.numeric(year) == 2001 ~ "Reduced quantity ended",
                                       treatment == "altered" & as.numeric(year) >= 2003 ~ "Heat lamps added to two random subplots",
                                       treatment == "control" ~ "Unsheltered controls to account for effects from shelter infrastructure"))


# factors we need: year, site, plot, higher_order_organization, temperature, 
# precipitation, nutrients_added, nitrogen_amount, disturbance, grazing, 
# fire_frequency, time_since_fire, treatment comment, diversity_manipulated,
# plot_biomass, measurement_scale_biomass, richness, evenness, measurement_scale_cover


##### Community composition #####
ambient_ramps <- c(3, 4, 6, 7, 11, 15)
altered_ramps <- c(2,5,9, 10, 12, 13)
control_ramps <- c(1, 8, 14)

RaMPs_PlantCommunityComposition <- ramps %>% 
  mutate(A_g_m2 = A*10,
         B_g_m2 = B*10,
         C_g_m2 = C*10,
         D_g_m2 = D*10) %>% 
  dplyr::select(RecType,RecYear,season,RampNo,Species,A_g_m2,B_g_m2,C_g_m2,D_g_m2) %>% 
  dplyr::rename(A = A_g_m2,
                B = B_g_m2,
                C = C_g_m2,
                D = D_g_m2) %>% 
  tidyr::gather(key="plot", value="plot_biomass_spp", A, B, C, D) %>% 
  merge(., RaMPs_ANPP_working, 
        by=c("RecYear", "season", "RampNo", "plot")) %>% 
  mutate(Pseudo_PercCover = plot_biomass_spp/plot_biomass*100) %>% 
  dplyr::rename(year = RecYear,
                subplot = plot) %>% 
  mutate(site = "KNZ",
         ramps = "RampNo",
         experiment = "RaMPs",
         treatment  = case_when(RampNo %in% ambient_ramps ~ "ambient",
                                RampNo %in% altered_ramps ~ "altered",
                                RampNo %in% control_ramps ~ "control"),
         block = case_when(RampNo %in% c(1:5) ~ "Block1",
                           RampNo %in% c(6:10) ~ "Block2",
                           RampNo %in% c(11:15) ~ "Block3")) %>% 
  filter(year != "1997") %>% # 1997 was year before shelters were put up
  #group_by(year, season, RampNo, site, ramps, experiment, treatment) %>% 
  #summarize(plot_biomass = mean(plot_biomass)) %>% 
  tidyr::unite("higher_order_organization", c(experiment, block), remove = F) %>% #ramps, RampNo
  tidyr::unite("uniqueID", c(higher_order_organization, ramps, RampNo, subplot), remove = F) %>%
  dplyr::rename(plot = RampNo) %>% 
  dplyr::select(year, season, site, plot, higher_order_organization, uniqueID, treatment, Pseudo_PercCover, Species) %>% 
  mutate(nutrients_added = "no_fertilizer",
         nitrogen_added = 0,
         disturbance = "undisturbed",
         grazing = "ungrazed",
         fire_frequency = 1,
         time_since_fire = NA,
         diversity_manipulated = "naturally_assembled",
         measurement_scale_biomass = NA,
         richness = NA,
         evenness = NA,
         measurement_scale_cover = NA) %>% 
  mutate(treatment_comment = case_when(treatment == "altered" & as.numeric(year) <= 2001 ~ "Could be reduced quantity, altered pattern, or both",
                                       treatment == "altered" & as.numeric(year) == 2001 ~ "Reduced quantity ended",
                                       treatment == "altered" & as.numeric(year) >= 2003 ~ "Heat lamps added to two random subplots",
                                       treatment == "control" ~ "Unsheltered controls to account for effects from shelter infrastructure"))

```


# Cleaning data A1 & A2 (ConSME)- Maya
Note: for ConSME, it looks like there are only 4 years of species comp data, and 2 years of biomass data. There should be more biomass data for sure though. 
```{r}

#Reading in the data
df_A1_KNZ <- read.csv(file.path(L0_dir, "KNZ/A1_CME011_Sppcomp.csv"), stringsAsFactors = FALSE)
df_A2_KNZ <- read.csv(file.path(L0_dir, "KNZ/A2_CME012_Biomass.csv"), stringsAsFactors = FALSE)
df_A3_KNZ <- read.csv(file.path(L0_dir, "KNZ/A3_CME01_treatments.csv"), stringsAsFactors = FALSE)

#I added the Konza species list to the drive as well
spplist <- read.csv(file.path(L0_dir, "KNZ/KNZ_specieslist_PPS011.csv"), stringsAsFactors = FALSE)


# Initial checks 
unique(df_A1_KNZ$RecYear) ##There are only 4 years of recorded species comp data: 2019-2022.  
unique(df_A2_KNZ$Recyear) ##There are only 2 years of recorded biomass data: 2020 & 2021.
unique(df_A1_KNZ$Watershed) ##2 watersheds: N1A & N4B
unique(df_A2_KNZ$Watershed) ##Same two watersheds: N1A & N4B
unique(sort(df_A1_KNZ$Block)) ##Blocks A-R
unique(sort(df_A2_KNZ$Block)) ##Same blocks A-R
unique(sort(df_A1_KNZ$Taxa)) ## Some non-plant observations that need to be removed; 


### A1 - sppcomp

## I am going to go ahead and remove non-plant things, and make invalid (negative) entries into 0's (NOT SURE IF THIS IS THE RIGHT PROTOCOL)
df_A1_KNZ.clean <- subset(df_A1_KNZ, !(Taxa %in% c("NA NA", "Dung ", "Rock ", 
                                                   "Litter", "Litter ", "Unknown seedlings")))
df_A1_KNZ.clean$Cover <- replace(df_A1_KNZ.clean$Cover, df_A1_KNZ.clean$Cover < 0, 0) 


## Looking through the dataset, I noticed some species names are different/misspelled and some entries are labelled with the wrong species number, so I am going to double check some species numbers, then just merge with the Konza species list to take care of the misspellings.

for(i in 1:nrow(df_A1_KNZ.clean)){
  if(df_A1_KNZ.clean$Taxa[i] == "Symphyotrichym (Aster) ericoides"){df_A1_KNZ.clean$Sppnum[i] <- 58}
  if(df_A1_KNZ.clean$Taxa[i] == "Symphyotrichym (Aster) oblongifolium"){df_A1_KNZ.clean$Sppnum[i] <- 59}
}


#Taking just the columns I want (for now) from the species list
spplist_short <- spplist %>% 
  unite(species, genus:species, sep = "_" ) %>%
  mutate(Sppnum = code) %>%
  dplyr::select(Sppnum, species) %>%
  add_case(Sppnum =c(-999, 0), 
           species = c("unknown","bare_ground"))

#Joining the species list and data together, so now we have correct spellings of species.
df_A1_KNZ.clean <- left_join(df_A1_KNZ.clean, spplist_short, by = "Sppnum")


#Rename columns to match datasheet outline
df_A1_KNZ.clean <- df_A1_KNZ.clean %>%
  mutate(site = as.factor("KNZ"),
         rectype = RecType,
         datacode = Datacode,
         year = as.numeric(RecYear),
         recdate = RecDate,
         season = as.factor(Season),
         watershed = as.factor(Watershed),
         block = Block,
         plot = Plot,
         sppcode = Sppnum,
         species = species,
         abundance = Cover,
         comments = Comments,
         experiment  = "ConSME",
         .keep = "unused") %>%
  unite(higher_order_organization, 
        experiment, site, watershed, 
        block, sep = "_", remove = FALSE) %>%
  unite(uniqueid, higher_order_organization, plot, sep = "_", remove = FALSE) %>%
  mutate(higher_order_organization = as.factor(higher_order_organization),
         uniqueid = as.factor(uniqueid),
         measurement_scale_cover = "1m2")

#Keeping the highest plot/species cover per season by summarizing data by the max abundance for every variable but season... there is probably a better way to do this lol.
df_A1_KNZ.clean <- df_A1_KNZ.clean %>% group_by(datacode, experiment, year, 
                                                 recdate, higher_order_organization, uniqueid, 
                                                 site, watershed, block,
                                                 plot, sppcode, species, Taxa, measurement_scale_cover, comments) %>% 
  summarize(abundance = max(abundance))


## Calculating relative abundance
df_A1_KNZ.clean <- df_A1_KNZ.clean %>%
 group_by(year, plot, uniqueid, watershed)%>%
  mutate(total_abundance = sum(abundance))%>%
  group_by(year, plot, uniqueid, watershed, species)%>%
  mutate(relative_abundance = (abundance/total_abundance) * 100)%>%
  ungroup()

## Calculating richness and evenness
A1_richness <- df_A1_KNZ.clean %>%
  community_diversity(time.var = "year",
                      abundance.var = "relative_abundance",
                      replicate.var = "uniqueid", metric = "Shannon")

A1_diversity <- df_A1_KNZ.clean %>% 
  #calculates richness and evenness as Evar
  community_structure(time.var = "year", 
                      abundance.var = "relative_abundance",
                      replicate.var = "uniqueid",
                      metric = "SimpsonEvenness") %>%
  left_join(A1_richness, by = c("year","uniqueid"))


#Creating an dataframe for all the distinct descriptive columns
df_A1_KNZ.env <- df_A1_KNZ.clean %>% 
    distinct(datacode, experiment, year, 
             higher_order_organization, uniqueid, 
             site, watershed, block, plot)

#Joining the diversity dataframe to the distinct columns dataframe
A1_diversity <- left_join(A1_diversity, df_A1_KNZ.env, by = c("year", "uniqueid"), keep = FALSE) %>%
  mutate(plot = as.factor(plot))
    


### A2-biomass

#Renaming variables so they match the sppcomp data (and so that they're all lowercase because I'm weird and I need things to match)
df_A2_KNZ.clean <- df_A2_KNZ %>%
  rename(datacode = DataCode,
         rectype = RecType,
         year = Recyear,
         watershed = Watershed,
         block = Block,
         plot = Plot,
         strip = Strip,
         lvgrass = Lvgrass,
         forb = Forb,
         woody = Woody,
         pdead = Pdead,
         comments = Comments
         ) %>% 
  mutate(site = "KNZ",
         experiment  = "ConSME") %>%
  unite(higher_order_organization, 
        experiment, site, watershed, 
        block, sep = "_", remove = FALSE) %>%
  unite(uniqueid, higher_order_organization, plot, sep = "_", remove = FALSE) %>%
  mutate(higher_order_organization = as.factor(higher_order_organization),
         uniqueid = as.factor(uniqueid))

#Replacing values that were recorded as -999 to 0 (NOT SURE IF THIS IS THE RIGHT PROTOCOL)
df_A2_KNZ.clean$lvgrass <- replace(df_A2_KNZ.clean$lvgrass, df_A2_KNZ.clean$lvgrass < 0, 0) 
df_A2_KNZ.clean$forb <- replace(df_A2_KNZ.clean$forb, df_A2_KNZ.clean$forb < 0, 0)
df_A2_KNZ.clean$woody <- replace(df_A2_KNZ.clean$woody, df_A2_KNZ.clean$woody < 0, 0)
df_A2_KNZ.clean$pdead <- replace(df_A2_KNZ.clean$pdead, df_A2_KNZ.clean$pdead < 0, 0)

#Since biomass was taken in 2 strips (both 1m2) per plot, I'm going to average the two strips for each plot... not sure if this is what we want... unfortunately, this gets rid of comments 
df_A2_KNZ.clean <- df_A2_KNZ.clean %>% group_by(datacode, year, higher_order_organization, uniqueid, watershed, block, plot) %>%
  summarize(lvgrass_avg = mean(lvgrass),
            forb_avg = mean(forb),
            woody_avg = mean(woody),
            pdead_avg = mean(pdead)) %>% 
  ungroup() %>%
  group_by(year, watershed, uniqueid, plot) %>% # Summing the biomass together -- does not include pdead
  mutate(plot_biomass = sum(lvgrass_avg, forb_avg, woody_avg),
         plot = as.factor(plot)) %>%
  ungroup()




# Putting A1_diversity and A2 together
df_A1_A2_KNZ.clean <- left_join(A1_diversity, df_A2_KNZ.clean,
                                by = c("year", "higher_order_organization", "uniqueid", 
                                       "watershed", "block", "plot"), keep = FALSE) %>%
  mutate(measurement_scale_cover = "1m2",
         measurement_scale_biomass = "0.1m2") %>%
  rename(source = datacode.x)

#END RESULTs: 
view(df_A1_KNZ.clean)
view(df_A1_A2_KNZ.clean)






#Taking the extra info and putting it into a new metadata data frame?
A1_A2_metadata <- df_A1_KNZ.env %>% mutate(fire_frequency = NA,
                                              grazing = "ungrazed",
                                              time_since_fire = NA,
                                           nutrients_added = NA,
                                           nitrogen_amount = NA,
                                           disturbance = NA) #I am unsure what we are counting as disturbance... like is bison grazing considered a disturbance?

#Choosing just the columns I want from the treatment dataset
df_A3_KNZ.short <- df_A3_KNZ %>% select(watershed, block, plot, trt)


#Joining the species comp data and treatment datasets
A1_A2_metadata <- left_join(A1_A2_metadata, df_A3_KNZ.short, by = c("watershed", "block", "plot")) %>%
  mutate(treatment = factor(trt),
         watershed = as.factor(watershed),
         .keep = "unused",
         treatment_comment = NA)

#NOTE: for treatment, the presence of an "B" mean bison were allowed on the plot, "S" means small mammals had access to the plot, and "I" means insects had access. The "X" in place of any of these letters means the herbivore was excluded.

#Adding in the fire and grazing info
for(i in 1:nrow(A1_A2_metadata)){
  if(A1_A2_metadata$watershed[i] == "N1A"){A1_A2_metadata$fire_frequency[i] <- 1}
  if(A1_A2_metadata$watershed[i] == "N4B"){A1_A2_metadata$fire_frequency[i] <- 4}
  if(A1_A2_metadata$watershed[i] == "N1A"){A1_A2_metadata$time_since_fire[i] <- 0}
  if(A1_A2_metadata$watershed[i] == "N4B" & A1_A2_metadata$year[i] == 2019){A1_A2_metadata$time_since_fire[i] <- 0}
  if(A1_A2_metadata$watershed[i] == "N4B" & A1_A2_metadata$year[i] == 2020){A1_A2_metadata$time_since_fire[i] <- 1} #I should put code to calculate this instead of writing it by hand so that it is more integrative, but there are only two watersheds and 4 years of data and I just wanna do it the dumb way for now
  if(A1_A2_metadata$watershed[i] == "N4B" & A1_A2_metadata$year[i] == 2021){A1_A2_metadata$time_since_fire[i] <- 2}
  if(A1_A2_metadata$watershed[i] == "N4B" & A1_A2_metadata$year[i] == 2022){A1_A2_metadata$time_since_fire[i] <- 3}
  if(A1_A2_metadata$treatment[i] == "BSI"){A1_A2_metadata$grazing[i] <- "grazed"}
  if(A1_A2_metadata$treatment[i] == "BSX"){A1_A2_metadata$grazing[i] <- "grazed"}
}


#I have not yet done the precip and temp data


```
